Final draft

### 2021-07-12 final draft.Rmd
i
Adding section on Comparing Constant and Stochastic Volatility Models of Equity Returns in final draft
piccole modifiche a header

---
header-includes:
- \usepackage{txfonts}
- \usepackage{amsmath}
- \usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
- \usepackage{graphicx}
- \usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
- \usepackage{hyperref} %set the hyperlink to cite sections, equations...
- \urlstyle{same}
- \usepackage{comment}
- \usepackage{enumitem}
- \usepackage{ntheorem}
- \usepackage{lipsum}
- \usepackage{amssymb}
- \usepackage{xcolor}
- \theoremstyle{break}
- \newtheorem{thm}{Theorem}% theorem counter resets every \subsection
- \renewcommand{\thethm}{\arabic{thm}}
- \newtheorem{proposition}{Proposition}
- \newtheorem{assumption}{Assumption}
- \newtheorem{lemma}{Lemma}
- \theoremstyle{nonumberplain}
- \newtheorem{proof*}{Proof.}
- \newtheorem{definition}{Definition}
- \renewcommand{\P}{\mathbb{P}}
- \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
- \def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
- \setcounter{MaxMatrixCols}{10}
- \setlength{\oddsidemargin}{.0in}
- \setlength{\textwidth}{6 in}
- \setlength{\topmargin}{-.1in}
- \setlength{\textheight}{8.1in}
- \renewcommand{\baselinestretch}{1.5}
- \usepackage{amssymb}
- \newcommand{\E}{\mathbb E}
- \newcommand{\Q}{\mathbb Q}
- \newcommand{\M}{\mathbb M}
- \newcommand{\R}{\mathbb R}
- \newcommand{\B}{\mathcal B}
- \newcommand{\X}{\mathcal X}
- \newcommand{\Pd}{\mathbb P}
- \def\code#1{\texttt{#1}}
- \newtheorem{algorithm}{Algorithm}[section]

@@

\chapter{Markov Processes and State-Space Models}
\section{Notation}
Let us fix the basic notation that will be employed henceforth. Upper case letters, e.g. $X_t, Y_t,$ denote random variables, while lower case letters 
denote the realizations. Finite sequences are made compact using the semi-colon notation, e.g. $x_{0:t}=(x_0, x_1, \dots, x_t)$ for some $t \geq 0.$ The sets of outcomes are denoted by upper case calligraphic letters, e.g. $\mathcal{X}, \mathcal{Y}.$  The $\sigma$-algebra
of a set $\mathcal{X}$ reads $\mathcal{B}(\mathcal{X}).$ The pair $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ is a measurable space.
The densities are represented with lower case cursive letters. For example, the density of $X_t$ in $x_t$ is $p_t^\theta(x_t),$ where $\theta\in\Theta$ is some (eventually known) vector, while the density 
of $X_t$ in $x_t,$ given that $X_{t-1}$ has previously taken value $x_{t-1},$ is $p_t^\theta(x_t\mid x_{t-1}).$  Note that $p_t^\theta(x_t\mid x_{t-1})$ is a member of the parametric family with parameters $(\theta, x_{t-1})$.
The probability distribution of a random variable $X_t$ is represented by $\mathbb{P}_t(dx_t),$ while the probability distribution of a sequence $X_{0:t}$ is represented by $\mathbb{P}_t(dx_{0:t}).$ The Dirac measure, i.e., the measure that assigns probability $1$ to the singleton $\{x\},$ is denoted by $\delta_x(dy)$.The expectation operator is denoted by $\mathbb{E},$ where the section of $\mathbb{E}$ at $\mathbb{Q},$ i.e., $\mathbb{E}_{\mathbb{Q}}$ denotes the operator computed with respect to the probability distribution $\mathbb{Q}.$ 
Let $\varphi$ be a function defined on the random variable $X\propto\mathbb{Q}.$ Then, we denote the expected value of $\varphi(X)$ with 
$$
    \mathbb{E}_{\mathbb{Q}}[\varphi(X)]=
    \int_{\mathbb{X}}\varphi(x)\mathbb{Q}(dx).
$$

\section{Markov Processes}
Let $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ and $(\mathcal{Y}, \mathcal{B}(\mathcal{Y}))$ be two (eventually equal) measurable spaces. 
\begin{definition}
A function $P(x,dy): (\mathcal{X}, \mathcal{B}(\mathcal{X}))\rightarrow [0,1]$ such that 
\begin{itemize}
    \item for all $x\in\mathcal{X}, P(x,\cdot)$ is a probability measure on $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})$, and 
    \item for all subsets $A\in\mathcal{B}(\mathcal{Y}),$ the map $x\mapsto P(x,A)$ is measurable in $\mathcal{B}(\mathcal{X})),$
\end{itemize}
is defined a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})).$
\end{definition}
Notice that for any random variables $X_1$ and $X_0=x_0,$ 
$$
    \mathbb{P}_1(dx_{0:1})=\mathbb{P}_0(dx_{0})P_1(x_0,dx_1),
$$
where $P_1(x_0,dx_1)$ is a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{X}, \mathcal{B}(\mathcal{X})).$ This remark, provides some intuition about the link between probability kernels and conditional probabilities. As a matter of fact, it is possible to show that for given $X_{0:1}$ with probability distribution $\mathbb{P}_1(dx_{0:1}),$ 
\begin{equation}\label{prob_kern_condit_prob}
    \mathbb{P}_1(X_1\in dx_1\mid X_0=x_0)=P_1(x_0,dx_1).
\end{equation}

Probability kernels can be used to define Markov processes. Before moving to the definition, we will introduce some important concepts. Recall that given two measures $\mathbb{M}$ and $\mathbb{Q}$ defined on a measurable space $(\mathcal{X}, \mathcal{B}(\mathcal{X})),$ $\mathbb{Q}$ is absolutely continuous with respect to $\mathbb{M},$ that is $\mathbb{Q}\ll\mathbb{M},$ if for all sets $A\in\mathcal{B}(\mathcal{X})$ such that $\mathbb{M}(A)=0,$ then it must also hold that $\mathbb{Q}(A)=0.$ If $\mathbb{Q}(A)/\mathbb{M}(A)$ is well-defined for all $A,$ then the Radon-Nikodym theorem guarantees that there exists a measurable function $w(x)\geq0$ such that 
$$
    w(x)=\frac{\mathbb{Q}(dx)}{\mathbb{M}(dx)}, 
$$
meaning that, 
$$
    \mathbb{Q}(A)=\int_\mathcal{X} w(x) \mathbb{M}(dx).
$$

>>>>

---
header-includes:
- \usepackage{txfonts}
- \usepackage{amsmath}
- \usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
- \usepackage{graphicx}
- \usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
- \usepackage{hyperref} %set the hyperlink to cite sections, equations...
- \urlstyle{same}
- \usepackage{comment}
- \usepackage{enumitem}
- \usepackage{ntheorem}
- \usepackage{lipsum}
- \usepackage{amssymb}
- \usepackage{xcolor}
- \theoremstyle{break}
- \newtheorem{thm}{Theorem}% theorem counter resets every \subsection
- \renewcommand{\thethm}{\arabic{thm}}
- \newtheorem{proposition}{Proposition}
- \newtheorem{assumption}{Assumption}
- \newtheorem{lemma}{Lemma}
- \newtheorem{definition}{Definition}
- \theoremstyle{nonumberplain}
- \newtheorem{proof*}{Proof.}
- \newtheorem{definition}{Definition}
- \renewcommand{\P}{\mathbb{P}}
- \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
- \def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
- \setcounter{MaxMatrixCols}{10}
- \setlength{\oddsidemargin}{.0in}
- \setlength{\textwidth}{6 in}
- \setlength{\topmargin}{-.1in}
- \setlength{\textheight}{8.1in}
- \renewcommand{\baselinestretch}{1.5}
- \usepackage{amssymb}
- \newcommand{\E}{\mathbb E}
- \newcommand{\Q}{\mathbb Q}
- \newcommand{\M}{\mathbb M}
- \newcommand{\R}{\mathbb R}
- \newcommand{\B}{\mathcal B}
- \newcommand{\X}{\mathcal X}
- \newcommand{\Pd}{\mathbb P}
- \def\code#1{\texttt{#1}}
- \newtheorem{algorithm}{Algorithm}[section]

@@

\chapter{Markov Processes and State-Space Models}
\section{Notation}
Let us fix the basic notation that will be employed henceforth. Upper case letters, e.g. $X_t, Y_t,$ denote random variables, while lower case letters 
denote the realizations. Finite sequences are made compact using the semi-colon notation, e.g. $x_{0:t}=(x_0, x_1, \dots, x_t)$ for some $t \geq 0.$ The sets of outcomes are denoted by upper case calligraphic letters, e.g. $\mathcal{X}, \mathcal{Y}.$  The $\sigma$-algebra
of a set $\mathcal{X}$ reads $\mathcal{B}(\mathcal{X}).$ The pair $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ is a measurable space.
The densities are represented with lower case cursive letters. For example, the density of $X_t$ in $x_t$ is $p_t^\theta(x_t),$ where $\theta\in\Theta$ is some (eventually known) vector, while the density 
of $X_t$ in $x_t,$ given that $X_{t-1}$ has previously taken value $x_{t-1},$ is $p_t^\theta(x_t\mid x_{t-1}).$  Note that $p_t^\theta(x_t\mid x_{t-1})$ is a member of the parametric family with parameters $(\theta, x_{t-1})$.
The probability distribution of a random variable $X_t$ is represented by $\mathbb{P}_t(dx_t),$ while the probability distribution of a sequence $X_{0:t}$ is represented by $\mathbb{P}_t(dx_{0:t}).$ The Dirac measure, i.e., the measure that assigns probability $1$ to the singleton $\{x\},$ is denoted by $\delta_x(dy)$.The expectation operator is denoted by $\mathbb{E},$ where the section of $\mathbb{E}$ at $\mathbb{Q},$ i.e., $\mathbb{E}_{\mathbb{Q}}$ denotes the operator computed with respect to the probability distribution $\mathbb{Q}.$ 
Let $\varphi$ be a function defined on the random variable $X\propto\mathbb{Q}.$ Then, we denote the expected value of $\varphi(X)$ with 
$$
    \mathbb{E}_{\mathbb{Q}}[\varphi(X)]=
    \int_{\mathcal{X}}\varphi(x)\mathbb{Q}(dx).
$$

\section{Markov Processes}
Let $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ and $(\mathcal{Y}, \mathcal{B}(\mathcal{Y}))$ be two (eventually equal) measurable spaces. 
\begin{definition}
A function $P(x,dy): (\mathcal{X}, \mathcal{B}(\mathcal{X}))\rightarrow [0,1]$ such that 
\begin{itemize}
    \item for all $x\in\mathcal{X}, P(x,\cdot)$ is a probability measure on $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})$, and 
    \item for all subsets $A\in\mathcal{B}(\mathcal{Y}),$ the map $x\mapsto P(x,A)$ is measurable in $\mathcal{B}(\mathcal{X})),$
\end{itemize}
is defined a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})).$
\end{definition}
Notice that for any random variables $X_1$ and $X_0=x_0,$ 
$$
    \mathbb{P}_1(dx_{0:1})=\mathbb{P}_0(dx_{0})P_1(x_0,dx_1),
$$
where $P_1(x_0,dx_1)$ is a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{X}, \mathcal{B}(\mathcal{X})).$ This remark, provides some intuition about the link between probability kernels and conditional probabilities. As a matter of fact, it is possible to show that for given $X_{0:1}$ with joint probability distribution $\mathbb{P}_1(dx_{0:1}),$ 
\begin{equation}\label{prob_kern_condit_prob}
    \mathbb{P}_1(X_1\in dx_1\mid X_0=x_0)=P_1(x_0,dx_1).
\end{equation}

Probability kernels can be used to define Markov processes. Before moving to the definition, we will introduce some important concepts. Recall that given two measures $\mathbb{M}$ and $\mathbb{Q}$ defined on a measurable space $(\mathcal{X}, \mathcal{B}(\mathcal{X})),$ $\mathbb{Q}$ is absolutely continuous with respect to $\mathbb{M},$ that is $\mathbb{Q}\ll\mathbb{M},$ if for all sets $A\in\mathcal{B}(\mathcal{X})$ such that $\mathbb{M}(A)=0,$ then it must also hold that $\mathbb{Q}(A)=0.$ If $\mathbb{Q}(A)/\mathbb{M}(A)$ is well-defined for all $A,$ then the Radon-Nikodym theorem guarantees that there exists a measurable function $w(x)\geq0$ such that 
$$
    w(x)=\frac{\mathbb{Q}(dx)}{\mathbb{M}(dx)}, 
$$
meaning that, 
$$
    \mathbb{Q}(A)=\int_A w(x) \mathbb{M}(dx).
$$
